Peak area control doc

### 2023-08-29 targeted/targeted_control.Rmd
Complete rerun from MW file download

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options("readr.num_columns" = 0)

@@

clean_stans <- read.csv("../metadata/clean_stans.csv")
name_conversion <- clean_stans %>%
  select(new=compound_name, old=name_old) %>%
  distinct()
file_metadata <- read.csv("../metadata/filled_file_metadata.csv")
```

@@

```{r falkorRFs}
falkor_rundate <- strptime("190715", "%y%m%d")
falkor_stans <- clean_stans %>%
  filter(date_added<=falkor_rundate)

@@

manual_neg <- read_delim("raw_data/Mesotransect/MSDial_neg_data/Area_0_20235251247.txt", skip = 4, delim = "\t") %>%
  filter(`Metabolite name`!="Unknown") %>%
  select(compound_name=`Metabolite name`, starts_with("180821")) %>%
  pivot_longer(cols=starts_with("180821"), names_to = "filename", values_to="area") %>%
  mutate(polarity="neg")
manual_pos <- read_delim("raw_data/Mesotransect/MSDial_pos_data/Area_2_2023525133.txt", skip = 4, delim = "\t") %>%
  filter(`Metabolite name`!="Unknown") %>%
  select(compound_name=`Metabolite name`, starts_with("180821")) %>%
  pivot_longer(cols=starts_with("180821"), names_to = "filename", values_to="area") %>%
  mutate(polarity="pos")
manual_ints <- rbind(manual_pos, manual_neg) %>%
  mutate(compound_name=str_replace_all(compound_name, "Ingalls_|w/o MS2:|Unsettled: ", "")) %>%
  mutate(filename = ifelse(str_detect(filename, "DDA"),
                           str_replace(filename, "Poo", "DDA"),
                           filename)) %>%
  mutate(filename = ifelse(str_detect(filename, "QC"),
                           str_replace(filename, "Poo", "QC"),
                           filename))

# Once again, filter only those that Raafay integrated
clean_ints_transect <- manual_ints %>%
  filter(!str_detect(filename, "DDA")) %>%
  filter(compound_name!="Methyl 3-methylthiopropionate") %>%
  mutate(compound_name=str_replace(compound_name, "Monophosphate, 15N5", "monophosphate, 15N5")) %>%
  left_join(name_conversion, by=c("compound_name"="old")) %>% 
  mutate(compound_name=ifelse(is.na(new), compound_name, new)) %>% 
  select(-new) %>% 
  left_join(manual_quality, by=c("compound_name", "polarity")) %>%
  filter(integrated=="Y" &in_samps%in%c("Y", "Maybe?", "Kinda?", "Barely", "Kinda")) %>%
  select(-integrated, -in_samps) %>%
  mutate(area=as.numeric(area)) %>%
  anti_join(distinct(., compound_name, polarity) %>% filter(duplicated(compound_name)))
```

@@

```{r transectRFs}
transect_rundate <- strptime("180821", "%y%m%d")
transect_stans <- clean_stans %>%
  filter(date_added<=transect_rundate)

rfs_transect <- transect_stans %>%
  left_join(clean_ints_transect, by="compound_name") %>%
  filter(str_detect(filename, "_Std_")) %>% 
  # filter(compound_name=="N6-Acetyl-L-lysine") %>%
  # filter(compound_name=="Dimethylsulfonioacetate") %>%
  # ggplot() + geom_col(aes(x=area,y=filename))
  filter(compound_type!="Internal Standard") %>%
  select(-compound_type) %>%
  mutate(mix=paste0("H2OInMatrix|", paste0(mix, "InMatrix"), "|", paste0(mix, "InH2O"))) %>%
  filter(str_detect(filename, mix)) %>%
  mutate(type=str_replace(str_extract(filename, mix), "Mix[12]", "Mix")) %>%
  select(-mix) %>% 
  group_by(compound_name, type) %>%
  summarise(mean_area=mean(area, na.rm = TRUE), conc_um=unique(conc_um)) %>%
  pivot_wider(names_from = type, values_from=mean_area) %>% 
  mutate(rf_h2o=MixInH2O/conc_um) %>%
  mutate(rf_matrix=(MixInMatrix - H2OInMatrix)/conc_um) %>%
  mutate(rf_ratio=rf_matrix/rf_h2o)

@@

@@

```{r centerSkylineImport}
manual_pos <- read.csv("raw_data/Mesocenters/Ingalls_Lab_QE_Transition Results_POS.csv") %>%
  mutate(polarity="pos")
manual_neg <- read.csv("raw_data/Mesocenters/Ingalls_Lab_QE_Transition Results_NEG.csv") %>%
  mutate(polarity="neg")
manual_ints <- rbind(manual_pos, manual_neg) %>%
  select(filename=Replicate.Name, compound_name=Precursor.Ion.Name, 
         area=Area, polarity) %>%
  mutate(filename = ifelse(str_detect(filename, "DDA"),
                           str_replace(filename, "Poo", "DDA"),
                           filename)) %>%
  mutate(filename = ifelse(str_detect(filename, "QC"),
                           str_replace(filename, "Poo", "QC"),
                           filename)) %>%
  mutate(compound_name=str_replace(
    compound_name, "Monophosphate, 15N5", "monophosphate, 15N5")) %>%
  left_join(name_conversion, by=c("compound_name"="old")) %>% 
  mutate(compound_name=new) %>% 
  select(-new) %>% 
  left_join(manual_quality, by=c("compound_name", "polarity"))

clean_ints_center <- manual_ints %>%
  filter(integrated=="Y"&in_samps%in%c("Y", "Maybe?", "Kinda?", "Barely", "Kinda")) %>%
  select(-integrated, -in_samps) %>% 
  mutate(area=as.numeric(area))
```

@@

```{r centerRFs}
# Slightly weird because Mesocenter data doesn't have H2OinMatrix samples
# Instead, use rf_ratio from Falkor to swap in
center_rundate <- strptime("180205", "%y%m%d")
center_stans <- clean_stans %>%
  filter(date_added<=center_rundate)

rfs_center <- center_stans %>% 
  left_join(clean_ints_center, by="compound_name") %>% 
  filter(str_detect(filename, "_Std_")) %>%
  filter(compound_type!="Internal Standard") %>%
  select(-compound_type) %>% 
  mutate(mix="InMatrix|InH2O") %>% 
  filter(str_detect(filename, mix)) %>%
  mutate(type=str_extract(filename, mix)) %>%
  select(-mix) %>% 
  group_by(compound_name, type) %>%
  summarize(mean_area=mean(area, na.rm = TRUE), conc_um=unique(conc_um)) %>%
  pivot_wider(names_from = type, values_from=mean_area) %>%
  mutate(rf_h2o=InH2O/conc_um) %>%
  left_join(rfs_falkor %>% select("rf_ratio", "compound_name"), by="compound_name")

@@

```{r collect}
## Write out just the raw areas
raw_area_data <- rbind(
  cbind(clean_ints_falkor, cruise="Falkor"),
  cbind(clean_ints_center, cruise="Mesocenter"),
  cbind(clean_ints_transect, cruise="Mesotransect")
) %>%
  left_join(clean_stans, by="compound_name")
write.csv(raw_area_data, file = "made_data/raw_area_data.csv", row.names = FALSE)

@@

## Write out quant data
quant_data <- bind_rows(falkor_data, mesotransect_data, mesocenter_data) %>%
  filter(compound_name!="Choline")
write.csv(quant_data, file = "made_data/longdata.csv", row.names = FALSE)
write_csv(quant_data, file = "all_peaks.csv")
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, purl = TRUE)
options("readr.num_columns" = 0)

@@

clean_stans <- read.csv("../metadata/clean_stans.csv")
name_conversion <- clean_stans %>%
  select(new=compound_name, old=name_old) %>%
  distinct()
date_filter <- clean_stans %>%
  select(compound_name=name_old, date_added) %>%
  distinct()
file_metadata <- read.csv("../metadata/filled_file_metadata.csv")
```

@@

```{r falkorRFs}
falkor_rundate <- 190715
falkor_stans <- clean_stans %>%
  filter(date_added<=falkor_rundate)

@@

manual_neg <- read_delim("raw_data/Mesotransect/MSDial_neg_data/Area_0_20235251247.txt", skip = 4, delim = "\t") %>%
  filter(`Metabolite name`!="Unknown") %>%
  select(compound_name=`Metabolite name`, starts_with("180821")) %>%
  pivot_longer(cols=starts_with("180821"), names_to = "filename", values_to="area") %>%
  mutate(polarity="neg")
manual_pos <- read_delim("raw_data/Mesotransect/MSDial_pos_data/Area_2_2023525133.txt", skip = 4, delim = "\t") %>%
  filter(`Metabolite name`!="Unknown") %>%
  select(compound_name=`Metabolite name`, starts_with("180821")) %>%
  pivot_longer(cols=starts_with("180821"), names_to = "filename", values_to="area") %>%
  mutate(polarity="pos")
manual_ints <- rbind(manual_pos, manual_neg) %>%
  mutate(compound_name=str_replace_all(compound_name, "Ingalls_|w/o MS2:|Unsettled: ", "")) %>%
  mutate(filename = ifelse(str_detect(filename, "DDA"),
                           str_replace(filename, "Poo", "DDA"),
                           filename)) %>%
  mutate(filename = ifelse(str_detect(filename, "QC"),
                           str_replace(filename, "Poo", "QC"),
                           filename)) %>%
  anti_join(date_filter %>% filter(date_added>180821))

# Once again, filter only those that Raafay integrated
clean_ints_transect <- manual_ints %>%
  filter(!str_detect(filename, "DDA")) %>%
  filter(compound_name!="Methyl 3-methylthiopropionate") %>%
  filter(compound_name!="N(e)-Acetyl-Lysine") %>%
  filter(compound_name!="Valine") %>%
  mutate(compound_name=str_replace(compound_name, "Monophosphate, 15N5", "monophosphate, 15N5")) %>%
  left_join(name_conversion, by=c("compound_name"="old")) %>% 
  mutate(compound_name=ifelse(is.na(new), compound_name, new)) %>% 
  select(-new) %>% 
  left_join(manual_quality, by=c("compound_name", "polarity")) %>%
  filter(integrated=="Y" &in_samps%in%c("Y", "Maybe?", "Kinda?", "Barely", "Kinda")) %>%
  select(-integrated, -in_samps) %>%
  mutate(area=as.numeric(area)) %>%
  anti_join(distinct(., compound_name, polarity) %>% filter(duplicated(compound_name)))

# Fix sample 180821_Smp_MS6C315m_C which didn't get extraction standards added
ext_stans_list <- c(
  "Isethionic acid, 13C2",
  "L-Cysteic acid, 2H3",
  "Sulfoacetic acid, 13C2",
  "3-Sulfolactate, 13C3",
  "Taurine, 2H4"
)
update_df <- clean_ints_transect %>%
  filter(str_detect(filename, "Smp")) %>%
  filter(polarity=="neg") %>%
  filter(compound_name%in%ext_stans_list) %>%
  group_by(compound_name) %>%
  summarise(area=median(area)) %>%
  mutate(filename="180821_Smp_MS6C315m_C")
clean_ints_transect <- clean_ints_transect %>%
  rows_update(update_df, by = c("compound_name", "filename"))
```

@@

```{r transectRFs}
transect_rundate <- strptime("180821", "%y%m%d")
transect_stans <- clean_stans %>%
  filter(strptime(date_added, "%y%m%d")<=transect_rundate)

rfs_transect <- transect_stans %>%
  left_join(clean_ints_transect, by=c("compound_name", "polarity")) %>%
  filter(str_detect(filename, "_Std_")) %>% 
  # filter(compound_name=="N6-Acetyl-L-lysine") %>%
  # filter(compound_name=="Dimethylsulfonioacetate") %>%
  # ggplot() + geom_col(aes(x=area,y=filename))
  filter(compound_type!="Internal Standard") %>%
  select(-compound_type) %>%
  mutate(mix=paste0("H2OInMatrix|", paste0(mix, "InMatrix"), "|", paste0(mix, "InH2O"))) %>%
  filter(str_detect(filename, mix)) %>%
  mutate(type=str_replace(str_extract(filename, mix), "Mix[12]", "Mix")) %>%
  select(-mix) %>% 
  group_by(compound_name, type) %>%
  summarise(mean_area=mean(area, na.rm = TRUE), conc_um=unique(conc_um)) %>%
  pivot_wider(names_from = type, values_from=mean_area) %>% 
  mutate(rf_h2o=MixInH2O/conc_um) %>%
  mutate(rf_matrix=(MixInMatrix - H2OInMatrix)/conc_um) %>%
  mutate(rf_ratio=rf_matrix/rf_h2o)

@@

```{r transect fix sucrose/trehalose integrations, eval=FALSE}
# Sucrose and trehalose got integrated together which is obnoxious because we 
# have IS for each of those so it'd be really nice to get those good numbers
# Re-integrating using RaMS

library(RaMS)
# msdata <- list.files("../tmzMLs/neg", full.names = TRUE, 
#                      pattern = "170706|180821|180205") %>% 
#   grabMSdata(verbosity = 0)
# tresuc_msdata <- msdata$MS1[mz%between%pmppm(341.10839, 10)]
# IS_msdata <- msdata$MS1[mz%between%pmppm(353.1497, 10)]

suc_data <- tresuc_msdata[rt%between%c(11.2, 12.1)] %>%
  group_by(filename) %>%
  summarize(area=sum(int))
suc_IS_data <- IS_msdata[rt%between%c(11.2, 12.1)] %>%
  group_by(filename) %>%
  summarize(IS_area=sum(int))
# ggplot(suc_IS_data) + geom_col(aes(x=filename, y=IS_area))
suc_nM <- left_join(suc_data, suc_IS_data) %>%
  mutate(nM=area/IS_area*5000 * 0.0004/10 * 2) %>%
  mutate(compound_name="Sucrose")

tre_data <- tresuc_msdata[rt%between%c(12.1, 12.9)] %>%
  group_by(filename) %>%
  summarize(area=sum(int))
tre_IS_data <- IS_msdata[rt%between%c(12.1, 12.9)] %>%
  group_by(filename) %>%
  summarize(IS_area=sum(int))
# ggplot(suc_IS_data) + geom_col(aes(x=filename, y=IS_area))
tre_nM <- left_join(tre_data, tre_IS_data) %>%
  mutate(nM=area/IS_area*5000 * 0.0004/10 * 2) %>%
  mutate(compound_name="Trehalose")

update_df <- bind_rows(tre_nM, suc_nM) %>%
  select(compound_name, nM, filename) %>%
  mutate(polarity="neg") %>%
  mutate(filename=str_remove(filename, ".tmzML"))
mesotransect_data <- mesotransect_data %>%
  rows_upsert(update_df, by=c("compound_name", "nM", "filename", "polarity"))
```

@@

```{r centerSkylineImport}
manual_pos <- read.csv("raw_data/Mesocenters/Ingalls_Lab_QE_Transition Results_POS.csv") %>%
  mutate(polarity="pos")
manual_neg <- read.csv("raw_data/Mesocenters/Ingalls_Lab_QE_Transition Results_NEG.csv") %>%
  mutate(polarity="neg")
manual_ints <- rbind(manual_pos, manual_neg) %>%
  select(filename=Replicate.Name, compound_name=Precursor.Ion.Name, 
         area=Area, polarity) %>%
  mutate(filename = ifelse(str_detect(filename, "DDA"),
                           str_replace(filename, "Poo", "DDA"),
                           filename)) %>%
  mutate(filename = ifelse(str_detect(filename, "QC"),
                           str_replace(filename, "Poo", "QC"),
                           filename)) %>%
  mutate(compound_name=str_replace(
    compound_name, "Monophosphate, 15N5", "monophosphate, 15N5")) %>%
  left_join(name_conversion, by=c("compound_name"="old")) %>% 
  mutate(compound_name=ifelse(is.na(new), compound_name, new)) %>% 
  select(-new) %>% 
  left_join(manual_quality, by=c("compound_name", "polarity"))

clean_ints_center <- manual_ints %>%
  filter(integrated=="Y"&in_samps%in%c("Y", "Maybe?", "Kinda?", "Barely", "Kinda")) %>%
  select(-integrated, -in_samps) %>% 
  mutate(area=as.numeric(area)) %>%
  filter(!compound_name%in%c("Turicine", "Hydroxyisoleucine"))
```

@@

```{r centerRFs}
# Slightly weird because Mesocenter data doesn't have H2OinMatrix samples
center_rundate <- strptime("180205", "%y%m%d")
center_stans <- clean_stans %>%
  filter(strptime(date_added, "%y%m%d")<=center_rundate)

custom_h2o_mat <- clean_ints_center %>%
  filter(str_detect(filename, "Poo_Full")) %>%
  group_by(compound_name) %>%
  summarise(newH2OInMatrix=mean(area)*(1-(62.94*100/80)/100))

rfs_center <- center_stans %>% 
  left_join(clean_ints_center, by=c("compound_name", "polarity")) %>%
  filter(str_detect(filename, "_Std_")) %>%
  filter(compound_type!="Internal Standard") %>%
  select(-compound_type) %>% 
  mutate(mix="InMatrix|InH2O") %>% 
  filter(str_detect(filename, mix)) %>%
  mutate(type=str_extract(filename, mix)) %>%
  select(-mix) %>% 
  group_by(compound_name, type) %>%
  summarize(mean_area=mean(area, na.rm = TRUE), conc_um=unique(conc_um)) %>%
  pivot_wider(names_from = type, values_from=mean_area) %>%
  left_join(custom_h2o_mat) %>%
  # left_join(rfs_falkor) %>%
  mutate(rf_h2o=InH2O/conc_um) %>%
  mutate(rf_matrix=(InMatrix - newH2OInMatrix)/conc_um) %>%
  mutate(rf_ratio=rf_matrix/rf_h2o)

@@

```{r collect}
## Write out just the raw areas
raw_area_data <- rbind(
  cbind(clean_ints_falkor, cruise="Falkor"),
  cbind(clean_ints_center, cruise="Mesocenter"),
  cbind(clean_ints_transect, cruise="Mesotransect")
) %>%
  left_join(clean_stans, by=c("compound_name", "polarity"))
write.csv(raw_area_data, file = "made_data/raw_area_data.csv", row.names = FALSE)

@@

## Write out quant data
quant_data <- bind_rows(falkor_data, mesotransect_data, mesocenter_data) %>%
  filter(compound_name!="Choline") %>%
  mutate(nM=ifelse(compound_name=="Homarine", nM/25, nM))
write.csv(quant_data, file = "made_data/longdata.csv", row.names = FALSE)
write_csv(quant_data, file = "all_peaks.csv")
```
